Watch Out for the Bully! Job Interference Study on Dragonfly Network

### 2016-03-26 draft.tex
add a missing sentence
made changes based on John's comments

The high-radix, low-diameter dragonfly topology network will be the choice of many computing facilities for building their next generation supercomputers. Preliminary study shows that random placement with adaptive routing should be the rule of thumb to utilize such powerful network, since random placement can uniformly distribut traffic over the network, and adaptive routing can alleviate congestion. However, in this paper, we made the observation that although random placement coupled with adaptive routing can reach load-balanced over the network when multiple jobs running concurrently, it can not guarantee the best performance for each single job. The performance improvement of communication intensive jobs comes with sacrifice of those less intensive jobs. We refer this phenomenon as ``bully" between concurrently running jobs. We analyze the detailed reason behind this ``bully" behavior between jobs from the underneath network level. Our study is based on a sophisticated, high-fidelity discrete even-driven simulation tool. We believe the observations and analysis in this work will make the chosen of job placement and routing strategies on dragonfly network more cautiously in future.

@@

The low latency and high bandwidth interconnect network plays a critical role in improving HPC system performance. As the computing ability continue to grow due to many-core nodes with accelerators, the network is now becoming a scare resources. HPC community start to explore the new network topology to replace the commonly used torus-connected network. The high-radix, low-diameter dragonfly topology can lower the cost of interconnect network, improve network bandwidth, reduce the packets traverse latency \cite{dally-dragonfly}. This make the dragonfly network an very promising choice for building Petaflop and beyond supercomputers. 
%The procurement results from Trinity and CORAL project shows the great performance improvement can be got from using dragonfly network.

While dragonfly topology provides low-latency and high-bandwidth interconnected network, how to efficiently utilize it is greatly relying on the upper level system software. In particular, intelligent job placement to avoid interference is of paramount importance for guarantee the performance of concurrently running jobs that sharing network resources \cite{bhatele2015, dskinner}. Some research work suggest that random job placement coupled with adaptive routing can alleviate local congestion, eliminate hot-spots and reach load-balance for dragonfly networks \cite{jain-sc14, bhatele-sc11, brandt2014}.

In this work, we study the interference between concurrently running jobs on dragonfly network with different job placement and routing schemes. We use CODES, a HPC network simulation toolkit with high fidelity, excellent scalability for packet level simulation \cite{codes}. And being unorthodox to most existing work that use synthetic workload, we use traces of three representative applications from DOE Design Forward Project \cite{designforwardwebpage}. The advantage of using real application traces is quite obvious, these traces can capture application's communication behavior in detail that usually omitted in synthetic workload for the purpose of simplicity. Applications usually are distinctive from each other in terms of data transfer amount, communication pattern, operation dependency etc. which are impossible to reproduce in synthetic workload.

We conduct extensive experiments with three applications running on dragonfly network with two job placement schemes and three routing schemes. We analyze the applications' performance in terms of communication time when they running with different placement\& routing configuration. Our evaluation is based on more than 1000 runs of simulation. Through extensive simulation with application traces, we make several interesting observations.

@@

    \item Random placement coupled with adaptive routing can not guarantee the performance of every job in the workload. The hot-spots free and load-balanced network status reached by using random placement with adaptive routing come with the cost of less communication-intensive jobs performance degradation.

    \item We identify the ``bully" behavior happening between concurrently running jobs on dragonfly network when random placement \& adaptive routing in use. Random placement \& adaptive routing redirect the traffic of communication intensive jobs to the routers through which less intensive jobs' traffic used to go, embezzle the network resource such as local and global channel bandwidth, slow down the communication of those less intensive jobs.

    \item We found that contiguous placement \& minimal routing can guarantee the consistent performance of each job running on dragonfly network, while may result in local network congestion and unbalanced network utilization. Contiguous placement \& minimal routing are preferred by less communication-intensive jobs, since it will prevent them from being ``bullied" by other communication-intensive jobs.

@@

These observations also reconfirm that interference between concurrently running jobs would make the performance of some jobs very unpredictable. The detailed analysis about the traffic and congestion in network level unveil the undercurrent beneath that interference.

To the best of our knowledge, using parallel application traces from production systems for the study of job interference on dragonfly network has not been reported so far. We believe the observations and analysis presented in this paper are valuable to a number of communities including HPC computing facility, system software development, and system administration. The computing facility should choose the configuration for dragonfly network implementation with consideration about resource sharing between concurrently running jobs. The workload manager would come up with better scheduling algorithm for jobs with distinctive communication behavior. And the system administration would be more efficient with the knowledge about jobs running status under different placement and routing configuration.

The rest of this paper is organized as follows. Section \ref{sec:background} describes the dragonfly network and the placement \& routing schemes. Section \ref{sec: methodology} talks about the use of CODES as research vehicle and three representative applications from DOE Design Forward Project. Section \ref{sec:workload-1} presents the analysis of three applications running on dragonfly network with different placement \& routing configuration. Section \ref{sec:workload-2} validates the observations we get in section \ref{sec:workload-1}. Section \ref{sec:related work} discusses related work. Finally, the conclusion is presented in Section \ref{sec:conclusion}.

@@

In this section, we talk about the Dragonfly topology and one of its implementation. We also present the placement and routing schemes  used on dragonfly network.

\subsection{Dragonfly Network}
\label{sec:network}
The dragonfly is a two-level hierarchical topology. It consists of several groups connected by All-to-All links. In each group, there could be ``a" number of routers connected via All-to-All local channels. For each router, there could be ``p" number of compute nodes attached to it via terminal links. The router has ``h" number of global channels that used as intergroup connections through which the router connected to other routers in different groups. Thus, the radix of each router is $k = a+h+p-1$. The implementation of such dragonfly network could be different between different computing facilities. The choosing of proper $a$, $h$, $p$ could be involving many factors, such as system scale, building cost and workload requirements.

For the purpose of loading balance, it is recommended that the proper dragonfly configuration should be like $a=2p=2h$\cite{kim-micro}. Under this configuration, the total number of groups ``g" in the dragonfly network would be $g = a*h+1 $, the total number of compute nodes ``N" in the network would be $N = p*a*g $. Figure \ref{fig:dragonfly-overview} illustrated an implementation of dragonfly network follows such configuration. There are 6 routers in each group, each router has 3 compute nodes attached to it. This dragonfly implementation consists of 19 groups and 342 nodes in total.

\begin{figure}[h!] 
  \centering
  \includegraphics[width=0.48\textwidth]{dragonfly-overview}
  \caption{Dragonfly Network Overview. Illustration of part of dragonfly network. There are 19 groups, 6 routers in each group, 3 compute nodes connected to each router, 342 nodes in total. Job\_1 requires for 6 nodes, get the allocation by Random placement. Job\_2 requires for 8 nodes, get the allocation by Contiguous placement. }
  \label{fig:dragonfly-overview}
\end{figure}

@@

For a parallel application requiring multiple compute nodes, job placement refers to the process of assigning the tasks of the application onto the system by a system software such as batch scheduler \cite{xu-cluster14}. In this work, we study two well cited placement policies for dragonfly systems:

@@

\textbf{Contiguous Placement:} In this scheme, the nodes will first be indexed consecutively in an increasing order of their group id and router id. The schedul assign each job required number of nodes by cutting chunks of nodes from the sequentially ordered list, as illustrated in Figure \ref{fig:dragonfly-overview} as blue nodes assigned to job\_2. The nodes been allocated to the same job will be adjacent to each other, usually they are attached to the same router or in the same group, depends on the size of the job. Contiguous placement confines the traffic of application into the local area, which may result in local network congestion and increase the possibility of hot-spots.

\textbf{Random Placement:} In this scheme, job will get a set of nodes that randomly selected from all the available nodes in the system. The nodes allocated to each job would be attached to different routers in different groups, as illustrated in Figure \ref{fig:dragonfly-overview} the red nodes assigned to job\_1. Some routers may be shared by different applications when using random placement. Random placement can spreading the traffic of application uniformly across the whole system, thus avoid the possible local congestion and reach load balance.

\subsection{Routing on Dragonfly}
\label{sec:routing-schemes}
The routing scheme refers to the strategy adopted to route messages from the source router to the destination router. The well studied routing schemes for dragonfly network include minimal routing, adaptive routing \cite{dally-dragonfly}, and progressive adaptive routing\cite{jiang}, and some of their variations \cite{won-prog-adaptive}. In this work we study three widely adopted routing mechanisms in dragonfly systems

\textbf{Minimal:} When the network deploy minimal routing, a message from source router will take the the shortest path to the destination router. When there are multiple shortest paths between the source and destination router, the message will be evenly divided among the paths. Minimal routing can guarantee the minimum hops the message takes from source to destination. However, it may result in congestion along the shortest path.

\textbf{Adaptive:} In this scheme, the path a message takes will be adaptively chosen between shortest and non-shortest path, depends on the congestion situation along that path. For the non-shortest path, an intermediate router will be randomly chosen. The message will first take the shortest path from source to that intermediate router, then it will take the shortest path from that intermediate router to the destination. Since adaptive scheme chooses routing path adaptively based on the congestion situation, it can void local congestion and reduce the possibility of hot-spots.

\textbf{Progressive Adaptive:} In this scheme, the decision to route minimally at each hop in the source group will be re-evaluated and any decision to route non-minimally at source router or at a subsequent hop is permant and is not re-evaluated \cite{jiang}. The progressive adaptive is capable of handling the case where a global link is congested but this information has not yet propagated back to the source router \cite{jiang}. The packet will take the non-minimal route only when the congestion is encountered.

@@

It is difficult to do experiment with concurrently running jobs on HPC system. One reason is that job placement and routing strategy are part of system configuration, it is impossible for user to make changes at will. Another reason is that it is unrealistic to reserve the system exclusively to run the same batch of jobs with desired placement and routing configurations and compare the results. Therefore, we resort to simulation in our work.

\subsection{Simulation Tool}
\label{sec:simulation-tool}
A simulation toolkit named CODES, Co-Design of Multilayer Exascale Storage Architectures, enables the exploration of simulating different HPC networks with high fidelity and great scalability \cite{codes}. CODES supports dragonfly network with packet-level high-fidelity simulation\cite{codes-dragonfly} \cite{misbah-tpds}. It can take real MPI application traces generated by SST DUMPI\cite{sst} to drive CODES network model. In this work, we focus on an in-depth analysis of interference  with different job placement \& routing schemes on dragonfly network.

@@

We choose three representative parallel applications from DOE Design Forward Project\cite{designforwardwebpage}. We believe these applications can be representative of a wide array of applications running on leadership-class supercomputers. Specifically, we study the Algebraic MultiGrid Solver (AMG), Geometric MultiGrid (MultiGrid) and CrystalRouter MiniApps.

\textbf{AMG:} The Algebraic MultiGrid Solver, is a parallel algebraic multi-grid solver for linear systems arising from problems on unstructured mesh physics packages. It has been derived directly from the BoomerAMG solver that is being developed in the Center for Applied Scientific Computing (CASC) at LLNL\cite{amg}. 
% The traces of AMG we chosen are comprised of 216 MPI ranks.

\textbf{MultiGrid:}  geometric multi-grid v-cycle from the production elliptic solver BoxLib, a software framework for massively parallel block-structured adaptive mesh refinement (AMR) codes\cite{boxlib}. It is widely used for structured grid physics packages. 
% The traces of MultiGrid we chosen are comprised of 125 MPI ranks.

\textbf{CrystalRouter:} the extracted communication kernel of the full application Nek5000\cite{nek5000}, which is a spectral element CFD application developed at Argonne National Laboratory\cite{crystalrouter}. It features spectral element multi-grid solvers coupled with a highly scalable, parallel coarse-grid solver that is widely used for projects including ocean current modeling, thermal hydraulics of reactor cores,and spatiotemporal chaos. 
% The traces of CrystalRouter we chosen are comprised of 100 MPI ranks.

@@

\begin{table}[ht]
\begin{center}
\caption{The notation for different placement \& routing configurations} 
\label{tab: placement routing configs}
%\centering % Centers the table on the page, comment out to left-justify
\begin{tabular}{l c c c }
\toprule % Top horizontal line
\toprule
&\multicolumn{3}{c}{Routing} \\ % Amalgamating several columns into one cell is done using the \multicolumn command as seen on this line
\cmidrule(l){2-4}
Placement  & Minimal & Adaptive & Progressive adaptive\\ % Column names row
\midrule % In-table horizontal line
Contiguous  &  CM   &   CA   &  CPA   \\ % Content row 1
\midrule
Random  &   RM  &   RA   &  RPA   \\

@@

The dragonfly network studied in this work consists of 33 groups, each group contains 8 routers and each router has 4 compute nodes attached to it. Overall, there are 264 routers and 1056 compute nodes in the network. The ratio between aggregrated local and global channel bandwidth are set propotional to the Cray Cascade system\cite{faanes} based on the dragonfly topology\cite{dally-dragonfly}.

In this work we simulate and analyze six combinations of placement policies and routing schemes, which are summarized in Table \ref{tab: placement routing configs}. Our analysis focuses on both overall network workload and individual application performance. For overall workload analysis, we focus on the time spent on communication by the workload, the network traffic throughput and busy time of links and channels. For individual application analysis, we study the time spent on communication by each MPI rank, traffic and busy time of each router that belongs to specific application. In our experiments, we collect and analyze the following data:

\begin{itemize}
    \item router terminal link traffic and busy time.
    \item router local channel traffic and busy time.
    \item router global channel traffic and busy time.
    \item communication time spent by each MPI rank. 
\end{itemize}

With respect to random placement, we conduct our experiment with a set of distinctive random allocation, which are intended to eliminate the possibility of variation.

Workload \Rmnum{1} comprises the original AMG, MultiGrid and CrystalRouter. As we see in Table \ref{tab:apps-detail}, AMG has the least amount of data transfer, thus contributes the least amount of traffic in Wrokload \Rmnum{1}.

Workload \Rmnum{2} comprises sAMG, MultiGrid and CrystalRouter. sAMG consists of synthetic traces that are same the with AMG's original traces in every aspect expect that the data transfer amount 100x greater than AMG. The traffic from sAMG is the dominant in Workload \Rmnum{2}.

\begin{table}[ht]
\begin{center}
\caption{Details of Application Traces.} 
\label{tab:apps-detail}
%\centering % Centers the table on the page, comment out to left-justify
\begin{tabular}{l c c c }
\toprule % Top horizontal line
\toprule
&\multicolumn{3}{c}{Application Details} \\ % Amalgamating several columns into one cell is done using the \multicolumn command as seen on this line
\cmidrule(l){2-4}
App Name & num. rank & Data per Rank & Total Data\\ % Column names row
\midrule % In-table horizontal line
AMG  &    216 &   0.6MB   &     130MB\\ % Content row 1
\midrule
MultiGrid  &    125 &   5MB   &     625MB\\ 
\midrule
CrystalRouter  &   100  &  35MB    &     3500MB\\ 
\midrule
sAMG  &    216 &   60MB   &     13000MB\\ % Content row 1
\midrule % In-table horizontal line
\bottomrule % Bottom horizontal line
\end{tabular}
\end{center}
\end{table}

% \section{Study of Parallel Workload \Rmnum{1 }}
\section{Identify the ``Bully"}
\label{sec:workload-1}

% The first parallel workload analyzed in our experiment consists of three applications presented in section\ref{sec: application traces}. The details of this workload shown in Table \ref{tab: parallel workload-1}.

% 
% \begin{figure}[h!]
%   \centering
%   \includegraphics[width = 0.48\textwidth ]{data_amount}
%   \caption{Applications Data Transfer Amount. AMG has the least amount of data transfer, compared with MultiGrid and CrystalRouter.\NOTE{y axis in log?} }
%   \label{fig:3app-data-amount}
% \end{figure}
% 
% Figure \ref{fig:3app-data-amount} shows the data transfer amount of each MPI rank of three application. Based on this figure, we can tell that CrystalRouter is the most communication-intensive and AMG is the lest communication-intensive. 
% 
% 
% \subsection{Workload Analysis}
%  
% \begin{figure}[h!]
%   \centering
%   \includegraphics[width = 0.48\textwidth ]{wkld/wkld-commtime}
%   \caption{Time spent on communication by Workload I when 6 different placement and routing configurations are used. Random is better than contiguous in terms of improving communication efficiency. Communication time can be further reduced when (progressive) adaptive routing is used under both placement schemes. }
%   \label{fig:wkld-commtime}
% \end{figure}

\begin{table}[ht]
\begin{center}
\caption{Time spent on communication (MilliSecond) by Workload II when 6 different placement\&routing configuration in use.} 
\label{tab:wkld-commtime}
%\centering % Centers the table on the page, comment out to left-justify
\begin{tabular}{l c c c c c c }
\toprule % Top horizontal line
\toprule
&\multicolumn{6}{c}{Placement\&Routing Configuration} \\ % Amalgamating several columns into one cell is done using the \multicolumn command as seen on this line
\cmidrule(l){2-7}
	 & CM & CA & CPA & RM & RA & RPA \\ % Column names row
\midrule % In-table horizontal line
Runtime  &  1870 & 1250 & 1248 & 860 & 835 &830\\ % Content row 1

The communication time spent by parallel workload \Rmnum{1} under different placement and routing configuration shown in Table \ref{tab:wkld-commtime}. The contiguous placement coupled with minimal routing results in highest communication time. Progressive adaptive and adaptive routing have comparable results when coupled with contiguous placement. The random placement results the comparable performance with any routing schemes and they all outperform the contiguous placement configurations.

@@

\subsection{Workload Analysis}

>>>>

The high-radix, low-diameter dragonfly network will be the choice of many computing facilities for building their next generation supercomputers. Preliminary studies show that random placement with adaptive routing should be the rule of thumb to utilize such networks, since random placement with adaptive routing can uniformly distribute traffic over the network and alleviate congestion. However, in this paper, we made the observation that while random placement coupled with adaptive routing can load-balance the network when multiple jobs are running concurrently, they can not guarantee the best performance for each single job. The performance improvement of communication-intensive jobs comes with the sacrifice of less intensive jobs. We refer this phenomenon as ``bully" between concurrently running jobs. We presented a detailed analysis about this ``bully" behavior between jobs at the network level based on a sophisticated, high-fidelity discrete even-driven simulation tool. The observations and analysis presented in this work illuminate a path to the adoption of more efficient job placement policy and routing algorithm for HPC systems with dragonfly network.

@@

The low latency and high bandwidth interconnect network plays a critical role in HPC system performance. As computational capabilities continue to grow due to many-core nodes with accelerators, the network is increasingly becoming a scare resources. The high-radix, low-diameter dragonfly topology can lower the \NOTE{?} cost of the interconnect, improve network bandwidth and reduce packet latency \cite{dally-dragonfly}. This make the dragonfly network an very promising choice for building petaflop and beyond supercomputers. 
%The procurement results from Trinity and CORAL project shows the great performance improvement can be got from using dragonfly network.

While dragonfly topologies provide low-latency and high-bandwidth networking, efficiently utilizing it is a significant concern. In particular, intelligent job placement to avoid interference is of paramount importance to guarantee the performance of each job when it shares network resources with other concurrently running peers \cite{bhatele2015, dskinner}. Recent studies suggest that random job placement coupled with adaptive routing can alleviate local congestion, eliminate hot-spots and reach load-balance for dragonfly networks \cite{jain-sc14, bhatele-sc11, brandt2014}.

In this work, we study the interference between concurrently running jobs on dragonfly networks with different job placement and routing policies. We use CODES, an HPC network simulation toolkit with high fidelity and excellent scalability for packet level simulation \cite{codes}. Unlike much existing work that rely on synthetic workloads, we use traces of three representative applications from the DOE Design Forward Project \cite{designforwardwebpage}. The advantage of using real application traces is straight forward: \TODO{ these traces can capture application communication behavior in detail that usually omitted in synthetic workload for the purpose of simplicity}. Applications usually are distinctive from each other in terms of data transfer amount, communication pattern, operation dependency etc. which can be impractical to reproduce in a synthetic workload.

We conduct extensive experiments with three applications running on dragonfly network with two job placement policies and three routing policies. We analyze the applications' performance in terms of communication time when running with different placement and routing configurations. Our evaluation is based on more than 1000 runs of simulation. Through extensive simulation with application traces, we make several interesting observations.

@@

    \item Random placement coupled with adaptive routing can uniformly distribute traffic over the network, avoid hot-spots and alleviate traffic congestion. The network could reach the optimal performance by using random placement and adaptive routing.

    \item ``Bully" behavior is happening between concurrently running jobs on dragonfly network when random placement and adaptive routing in use. The performance improvement obtained by communication-intensive jobs come with the cost of less intensive jobs suffering performance degradation.

    \item Contiguous placement coupled with minimal routing can guarantee consistency performance of each job by prohibiting network resource sharing, though it results in local network congestion and unbalanced network utilization. Contiguous placement and minimal routing are preferred by less communication-intensive jobs, since it will prevent them from being ``bullied" by other communication-intensive jobs.

@@

%These observations also reconfirm that interference between concurrently running jobs would make the performance of some jobs very unpredictable. The detailed analysis about the traffic and congestion in network level unveil the undercurrent beneath that interference.

Based on our observations, the optimal job placement and routing solution for the workload running dragonfly system should be specific to job communication intensity. \emph{The random placement and adaptive routing are preferable to communication-intensive jobs, while the less intensive jobs should run with contiguous placement and minimal routing.} To the best of our knowledge, using parallel application traces from production systems for the study of job interference on dragonfly network has not been reported so far. We believe the observations and analysis presented in this paper are valuable to a number of communities including HPC computing centers, system software developers, and system administrators. The computing centers could choose the configuration for implementing a dragonfly network with consideration about resource sharing between concurrently running jobs. System software such as workload manager could come up with better scheduling algorithm for jobs with distinctive communication behavior. And system administrators could be able to make accurate predication about system availability with the knowledge of jobs running status under different placement and routing configurations.

The rest of this paper is organized as follows. Section \ref{sec:background} describes the dragonfly network and the placement and routing policies. Section \ref{sec: methodology} talks about the use of CODES as research vehicle and three representative applications from the DOE Design Forward Project. Section \ref{sec:workload-1} presents the analysis of three applications running on a dragonfly network with different placement and routing configurations. Section \ref{sec:workload-2} validates the observations we get in section \ref{sec:workload-1}. Section \ref{sec:related work} discusses the related work. Finally, the conclusion is presented in Section \ref{sec:conclusion}.

@@

In this section, we review the dragonfly topology, including the placement and routing policies examined in previous work.

\subsection{Dragonfly Network}
\label{sec:network}
The dragonfly is a two-level hierarchical topology. It consists of several groups connected by all-to-all links. In each group, there could be ``a" number of routers connected via all-to-all local channels. For each router, there could be ``p" number of compute nodes attached to it via terminal links. The router has ``h" number of global channels that used as intergroup connections through which the router connected to other routers in different groups. Thus, the radix of each router is $k = a+h+p-1$. The implementation of such dragonfly network could choose different value for ``a", ``p" and ``h" in different computing centers. The adoption of proper $a$, $h$, $p$ could be involving many factors, such as system scale, building cost and workload requirements.

For the purpose of load balance, it is recommended that the proper dragonfly configuration should follow $a=2p=2h$\cite{kim-micro}. Under this configuration, the total number of groups ``g" in the dragonfly network would be $g = a*h+1 $, the total number of compute nodes ``N" in the network would be $N = p*a*g $. An implementation of dragonfly network follows such configuration is illustrated in Figure \ref{fig:dragonfly-overview}. There are 6 routers in each group, each router has 3 compute nodes attached to it. This dragonfly implementation consists of 19 groups and 342 nodes in total.

\begin{figure}[h!] 
  \centering
  \includegraphics[width=0.48\textwidth]{dragonfly-overview}
  \caption{Dragonfly Network Overview. Illustration of part of dragonfly network. There are 19 groups, 6 routers in each group, 3 compute nodes connected to each router, 342 nodes in total. Job $J1$ requires for 7 nodes, gets its allocation by random placement. Job $J2$ requires for 8 nodes, gets its allocation by contiguous placement. }
  \label{fig:dragonfly-overview}
\end{figure}

@@

For a parallel application requiring multiple compute nodes, job placement policy refers to the way of assigning the tasks of the application onto the system by a system software such as batch scheduler \cite{xu-cluster14}. In this work, we study two alternative placement policies considered by the community for dragonfly systems:

@@

\textbf{Contiguous Placement:} In this policy, the compute nodes are assigned to job in a consecutive manner. The assignment fills up a group first, only crosses group boundaries and start to fill the next group as necessary. Job $J2$, illustrated as blue in Figure \ref{fig:dragonfly-overview}, gets the required amount of nodes by contiguous placement. Contiguous placement confines the tasks of a job into the same group, which may result in local network congestion and increase the possibility of hot-spots.

\textbf{Random Placement:} In this policy, job will get a set of nodes that randomly selected from all the available nodes in the system. The nodes assigned to each job would be attached to different routers in different groups, as illustrated in Figure \ref{fig:dragonfly-overview} the red nodes assigned to job $J1$. Some routers may be shared by different jobs when random placement in use. Random placement can distribute the tasks of  a job uniformly across the whole system, thus avoid the possible local congestion and reach load balance. However, random placement may cause possible congestion on global links.

\subsection{Routing on Dragonfly}
\label{sec:routing-schemes}
The routing policy refers to the strategy adopted to route messages \NOTE{(message refers to a stream of packets)} from the source router to the destination router. Previously studied routing policies for dragonfly network include minimal routing, adaptive routing \cite{dally-dragonfly}, and progressive adaptive routing\cite{jiang}, and some of their variations \cite{won-prog-adaptive}. In this work we study three widely adopted routing policies on dragonfly networks.

\textbf{Minimal:} In this policy, a message from source router will take the the shortest path to the destination router. When there are multiple shortest paths between the source and destination router, the message will be evenly divided among the paths. Minimal routing can guarantee the minimum hops the message takes from source to destination. However, it may result in congestion along the shortest path.

\textbf{Adaptive:} In this policy, the path a message takes will be adaptively chosen between shortest and non-shortest path, depends on the congestion situation along that path. For the non-shortest path, an intermediate router will be randomly chosen. The message will first take the shortest path from source to that intermediate router, then it will take the shortest path from that intermediate router to the destination. Since routing path are chosen adaptively based on the congestion situation, it can avoid local congestion and possible hot-spots. Adaptive routing collapses to minimal routing in the absence of traffic congestion.

\textbf{Progressive Adaptive:} In this policy, the decision to route minimally at each hop in the source group will be re-evaluated. And any decision to route non-minimally at source router or at a subsequent hop is permanent and won't be re-evaluated \cite{jiang}. The progressive adaptive is capable of handling the case where a global link is congested but this information has not yet propagated back to the source router \cite{jiang}. The packet will take the non-minimal route only when the congestion is encountered.

@@

It is difficult to do experiment with concurrently running jobs on HPC system\cite{zhou-ipdps-2015}\cite{jain-sc14}\cite{bhatele-sc11}\cite{jokanovic-ipdps-2015}. One reason is that job placement and routing policies are part of system configuration, it is impossible for user to make changes at will. Another reason is that it is unrealistic to reserve the system exclusively to run the same batch of jobs with desired placement and routing configurations and compare the results. Last but not least, configurable dragonfly networks that allow us to perform the exploration presented in this paper are hard to come by for the time being. Therefore, we resort to simulation in our work.

\subsection{Simulation Tool}
\label{sec:simulation-tool}
A simulation toolkit named CODES, Co-Design of Multilayer Exascale Storage Architectures, enables the exploration of simulating different HPC networks with high fidelity and great scalability \cite{codes}. CODES supports dragonfly network\cite{codes-dragonfly} \cite{misbah-tpds}, as well as other topologies such as torus\cite{misbah-pads-2014}\cite{ning-pads-2011} and fat-tree\cite{ning-pads-2015}, with packet-level high-fidelity simulation. It can drive these models through an MPI simulation later that utilizing DUMPI traces\cite{sst}. In this work, we perform an in-depth analysis of job interference with different job placement and routing policies on dragonfly network.

@@

We choose three representative parallel applications from the DOE Design Forward Project\cite{designforwardwebpage}. We believe these applications can be representative of a wide array of applications running on leadership-class supercomputers. Specifically, we study the \emph{Algebraic MultiGrid Solver} (AMG), \emph{Geometric Multigrid V-Cycle from production elliptic solver} (MultiGrid) and \emph{Crystal Router MiniApp} (CrystalRouter).

\textbf{AMG:} The Algebraic MultiGrid Solver is a parallel algebraic multi-grid solver for linear systems arising from problems on unstructured mesh physics packages. It has been derived directly from the BoomerAMG solver that is being developed in the Center for Applied Scientific Computing (CASC) at LLNL \cite{amg}.

\textbf{MultiGrid:} The geometric multi-grid v-cycle from the production elliptic solver BoxLib is a software framework for massively parallel block-structured adaptive mesh refinement (AMR) codes \cite{boxlib}. It is widely used for structured grid physics packages.

\textbf{CrystalRouter:} The CrystalRouter MiniApp is the communication kernel of the full application Nek5000 \cite{nek5000}, which is a spectral element CFD application developed at Argonne National Laboratory. It features spectral element multi-grid solvers coupled with a highly scalable, parallel coarse-grid solver that is widely used for projects including ocean current modeling, thermal hydraulics of reactor cores, and spatiotemporal chaos.

@@

\NOTE{justify network size}The dragonfly network studied in this work consists of 33 groups, each group contains 8 routers and each router has 4 compute nodes attached to it. Overall, there are 264 routers and 1056 compute nodes in the network. The ratio between aggregated local and global channel bandwidth are set propotional to the Cray Cascade system\cite{faanes} based on the dragonfly topology\cite{dally-dragonfly}.

In this work, we study both the network behavior and the job interference when dragonfly network is configured with 6 different job placement and routing policy combinations, which are summarized in Table \ref{tab: placement routing configs}.

\begin{table}[ht]
\begin{center}
\caption{The notation for different placement and routing configurations} 
\label{tab: placement routing configs}
%\centering % Centers the table on the page, comment out to left-justify
\begin{tabular}{l c c c }
\toprule % Top horizontal line
\toprule
&\multicolumn{3}{c}{Routing Policies} \\ % Amalgamating several columns into one cell is done using the \multicolumn command as seen on this line
\cmidrule(l){2-4}
Placement Policies & Minimal & Adaptive & Progressive Adaptive\\ % Column names row
\midrule % In-table horizontal line
Contiguous  &  CM   &   CA   &  CPA   \\ % Content row 1
\midrule
Random  &   RM  &   RA   &  RPA   \\ 
\midrule % In-table horizontal line
\bottomrule % Bottom horizontal line
\end{tabular}
\end{center}
\end{table}

%Our analysis focuses on both overall network workload and individual application performance. For overall workload analysis, we focus on the time spent on communication by the workload, the network traffic throughput and busy time of links and channels. For individual application analysis, we study the time spent on communication by each MPI rank, traffic and busy time of each router that belongs to specific application. In our experiments, we collect and analyze the following data: 
Our analysis focus on the following metrics:
\begin{itemize}

    \item \textbf{Network Traffic:} The traffic refers to the amount of data go through each router in the network. For each router, the traffic go through its terminal link, local channel and global channel are analyzed respectively.

    \item \textbf{Network Saturated Time:} The saturated time refers to the time period when  the buffer of certain port in the router is full. Saturated time can denote the congestion on certain port.

    \item \textbf{Communication Time:} The communication time of each MPI rank refers to the time spent by the rank to complete its message exchange with all other ranks. For each job, this metric refers to the time spent to complete the message exchange by all its ranks.

\NOTE{limitations:}Since our study is about network behavior and job interference during communication, the time spent by each MPI rank on computation will not be counted in our experiments.

With respect to random placement, we conduct our experiments with 50 sets of distinctive allocation generated by random placement. The corresponding experimental results are median chosen, which intended to eliminate the possibility of variation.

%In this work, we use two sets of parallel workloads to quantified study the network behavior as well as job interference when the dragonfly network equipped with different placement and routing configurations.

In this study, we use two sets of parallel workloads. Workload \Rmnum{1} consists of AMG, MultiGrid and CrystalRouter. As we see in Table \ref{tab:apps-detail}, AMG has the least amount of data transfer, thus contributes the least amount of traffic in Workload \Rmnum{1}. Workload \Rmnum{2} consists of sAMG, MultiGrid and CrystalRouter. sAMG is a synthetic version of AMG, which is the same as AMG in every aspect expect that the data transfer amount 100x greater than AMG. sAMG contributes the dominant amount of traffic in Workload \Rmnum{2}.

Workload \Rmnum{1} is used to identify the ``bully" behavior between concurrently running jobs on dragonfly network. Workload \Rmnum{2} is used to verify the results and conclusion obtained from the study about Workload \Rmnum{1}.

Table \ref{tab:apps-detail} summarizes the details of each application. It shows the number of MPI ranks, the average data transfer amount of each rank and the total data transfer amount for each application.

\begin{table}[ht]
\begin{center}
\caption{The details of each application.} 
\label{tab:apps-detail}
%\centering % Centers the table on the page, comment out to left-justify
\begin{tabular}{l c c c }
\toprule % Top horizontal line
\toprule
&\multicolumn{3}{c}{Application Details} \\ % Amalgamating several columns into one cell is done using the \multicolumn command as seen on this line
\cmidrule(l){2-4}
App Name & Num. Rank & Avg. Data/Rank & Total Data\\ % Column names row
\midrule % In-table horizontal line
AMG  &    216 &   0.6MB   &     130MB\\ % Content row 1
\midrule
MultiGrid  &    125 &   5MB   &     625MB\\ 
\midrule
CrystalRouter  &   100  &  35MB    &     3500MB\\ 
\midrule
sAMG  &    216 &   60MB   &     13000MB\\ % Content row 1
\midrule % In-table horizontal line
\bottomrule % Bottom horizontal line
\end{tabular}
\end{center}
\end{table}

\section{Study of Parallel Workload \Rmnum{1 }}
%\section{Identify the ``Bully"}
\label{sec:workload-1}

We study the network behavior as well as application interference under different placement and routing configurations from two perspectives. First, we analyze the network performance in terms of workload communication time. Then, we isolate each application from the workload, and analyze the interference by scrutinizing the traffic go through the routers that serving each application.

% \begin{figure}[h!]
%   \centering
%   \includegraphics[width = 0.48\textwidth ]{data_amount}
%   \caption{Applications Data Transfer Amount. AMG has the least amount of data transfer, compared with MultiGrid and CrystalRouter.\NOTE{y axis in log?} }
%   \label{fig:3app-data-amount}
% \end{figure}
% 
% Figure \ref{fig:3app-data-amount} shows the data transfer amount of each MPI rank of three application. Based on this figure, we can tell that CrystalRouter is the most communication-intensive and AMG is the lest communication-intensive.

\subsection{Network Performance Analysis}
%  
% \begin{figure}[h!]
%   \centering
%   \includegraphics[width = 0.48\textwidth ]{wkld/wkld-commtime}
%   \caption{Time spent on communication by Workload I when 6 different placement and routing configurations are used. Random is better than contiguous in terms of improving communication efficiency. Communication time can be further reduced when (progressive) adaptive routing is used under both placement schemes. }
%   \label{fig:wkld-commtime}
% \end{figure}

\begin{table}[ht]
\begin{center}
\caption{Time spent on communication by Workload I when dragonfly network equipped with 6 different placement and routing configurations.} 
\label{tab:wkld-commtime}
%\centering % Centers the table on the page, comment out to left-justify
\begin{tabular}{l c c c c c c }
\toprule % Top horizontal line
\toprule
&\multicolumn{6}{c}{Placement and Routing Configurations} \\ % Amalgamating several columns into one cell is done using the \multicolumn command as seen on this line
\cmidrule(l){2-7}
	 & CM & CA & CPA & RM & RA & RPA \\ % Column names row
\midrule % In-table horizontal line
Time(ms)  &  1870 & 1250 & 1248 & 860 & 835 &830\\ % Content row 1

@@

The communication time spent by Workload \Rmnum{1} under different placement and routing configurations shown in Table \ref{tab:wkld-commtime}. The contiguous placement coupled with minimal routing (CM) results in highest communication time. When contiguous placement coupled with progressive adaptive and adaptive routing (CA, CPA), Workload \Rmnum{1} spends  comparable amount of time to complete communication. The random placement policy performs comparably with all routing policies and outperforms all contiguous placement configurations.

@@

\subsection{Network Performance Analysis}
